---
title: "i-Plus User Guide"
format: html
author:
  - name: Amelia Babb
    corresponding: true
    email: ababb@smith.edu
  - name: Teddy Martin
    corresponding: true
    email: tkmartin@smith.edu
  - name: Mags McLaughlin
    corresponding: true
    email: mmclaughlin@smith.edu
  - name: Tillie Slosser
    corresponding: true
    email: tslosser@smith.edu
---

## Introduction

i-Plus is a tool developed by students in SDS 410: Capstone in Statistical & Data Sciences for Elmhurst United Middle School. This tool is integrated into Google Sheets using Google Apps Scripts and allows for easier and more customizable visualization of data provided by the i-Ready testing program. The steps to install i-Plus into Google Sheets containing student data from Elmhurst are detailed in the **Installing i-Plus** section of this user guide. Additionally, certain data management practices are required to keep i-Plus functioning properly. These steps are described in the **Data Management** section of this guide. The most common error that arises when running i-Plus is described in the **Common Errors** section of this guide.

## Installing i-Plus

1. Access Google Apps Scripts by navigating to **Extensions > Apps Script** within the sheet you want to add i-Plus to

![](\images\image1)

2. Access the library menu by choosing the plus button next to **Libraries** within the side menu of the Scripts window

![](\images\image2)

3. Locate the i-Plus library by pasting the following code into the **Script ID** and selecting look up

```{r}
#| eval: false

1os5M3_htGWveLyoZGitnTZD3SnwptfRWwul15TmaodITFAuO_6_M4-3w
```

![](\images\image3)

4. Leave the version settings as the default HEAD. Change the identifier name to “iPlus” making sure to maintain the same capitalization. Select Add.

```{r}
#| eval: false

iPlus
```

![](\images\image4)

5. In the Code.gs file, delete any text that was automatically created when creating the Apps Script project. Copy the code below and paste it into the Code.gs file. Select the save button to save your changes to the project.

![](\images\image6)

```{r}
#| eval: false

// DO NOT EDIT THIS EXCEPT TO ADD A CALL TO A FUNCTION
function onOpen() {
  const ui = SpreadsheetApp.getUi(); 
  // ------------------------------------ CREATE MENU ------------------------------------
  ui.createMenu('iPlus')
    .addItem('Refresh Data', 'joinSheets')
    .addSubMenu(ui.createMenu('Pre-built Visualizations')
          .addItem('Scatterplot: Attendance vs. Score Improvement', 'preBuiltScatterplot')
          .addItem('Boxplot: Race vs. Score Improvement', 'prebuiltBoxplot')
          .addItem('Jitterplot: Score Improvement by Math Teacher', 'drawTeacherScoreJitterPlot'))
    .addSubMenu(ui.createMenu('Build-Your-Own Visualizations')
          .addItem('Scatterplot', 'customScatterplot')
          .addItem('Boxplot', 'customBoxplot')
          .addItem('Jitterplot', 'customJitterplot'))
    .addToUi();

 // ------------------------------------ UPDATE IF NEEDED ------------------------------------
  // check if update to merged data is needed
  const props = PropertiesService.getDocumentProperties();
  const needsUpdate = props.getProperty("needsUpdate");
  console.log()
  if (needsUpdate === "true") {
    // Example: show a toast message
    SpreadsheetApp.getActiveSpreadsheet().toast("Updates are needed based on recent edits. Please wait until the next message", "WARNING: Updating `Merged Data`", 200);
    joinSheets();
    SpreadsheetApp.getActiveSpreadsheet().toast("Updates are finished. You may continue", "`Merged Data` update finished", 200);
}
}


function onEdit(e) {
  const range = e.range;
  const oldValue = e.oldValue;
  const newValue = range.getValue();

  if (oldValue !== newValue) {
    const props = PropertiesService.getDocumentProperties();
    props.setProperty("needsUpdate", "true");
  }
}
```

6. Select the run button to run i-Plus and integrate it into your Google Sheet. 

![](\images\image5)

## Data Management

When running i-Plus, certain column headers and sheet names are used to find the correct data to visualize. Sheet names are most important for accessing the correct data to visualize. The following sheet names must stay the same to maintain functionality of i-Plus:

- Elmhurst Data
- 6th Grade Math iReady
- 7th Grade Math iReady
- 8th Grade Math iReady


Column headers are most important for creating visualizations contained in the **Pre-Built Visualizations** menu within i-Plus. The following column headers must stay the same to maintain functionality of these visualizations:

- Race
- Attendance
- Math Teacher
- Initial Scale Score
- Current Scale Score

Column order is important for only one column. The column containing student ID number must be the first column in all of the following sheets to properly merge i-Ready testing data with Elmhurst demographic data:

- Elmhurst Data
- 6th Grade Math iReady
- 7th Grade Math iReady
- 8th Grade Math iReady

In future iterations of Google Sheets used with i-Plus, please ensure that these standard sheet names, column headers and column order are maintained to retain functionality of i-Plus. If the headers are renamed, custom-built visualizations will still run smoothly, but the pre-built visualizations will break. If the ID column is not kept as the first column, both pre-built and custom-built visualizations will break, and i-Plus will be non-functional.

## Common Errors

The most common error occurs when trying to join the i-Ready data across grades. This error may occur when the refresh data functions are interrupted and not allowed to run in entirety. If this error occurs, you will receive the following error message as a popup in the sheets view.

![](\images\image8)

To remedy this error, any transformations to the i-Ready sheets must be deleted. In the following sheets, delete any column with the header "Grade" and "Score Change":

- 6th Grade Math iReady
- 7th Grade Math iReady
- 8th Grade Math iReady

There may be duplicates of these columns, and any duplicates should also be deleted, as seen in the following image.

![](\images\image9)

